---
title: "Data Analysis Notebook"
output: html_notebook
---


```{r}
#Authors: Ian McElveen and Cecilia Gonzales
#Author Date: 6/9/25

#Purpose: The purpose of this notebook is to house all data set transformation, cleansing, visualization, statistical analysis, and note-taking for the 2025 CU Athletic Department Sports Science Internship Program

#LAST UPDATED: 6/12/2025

#Including helpful libraries
library(tidyverse)
library(readxl)
library(aod)
library(dplyr)
library(ggplot2)
library(lubridate)

#Example of importing our first data set from a CSV file to a data frame.

# df <- read.csv("./foo/bar.csv")
```



```{r load in datasets}
#Load in MBB - Statistic Tracking Report csv
stats <- read.csv("data-sets/data-sets-uncompressed/data-sets-compressed/Reactive Strength Index/MBB - Statistic Tracking Report.csv")

# Load in ACWR - Kinexon MBB
kinexon_ACWR <- read.csv("data-sets/data-sets-uncompressed/data-sets-compressed/Reactive Strength Index/ACWR - Kinexon MBB.csv")

#Loading in VALD - Dynamo MBB
VALD_dynamo <- read.csv("data-sets/data-sets-uncompressed/data-sets-compressed/Reactive Strength Index/VALD - Dynamo MBB.csv")

# Load in VALD - ForceDecks MBB csv
VALD_ForceDecks <- read.csv("data-sets/data-sets-uncompressed/data-sets-compressed/Reactive Strength Index/VALD - ForceDecks MBB.csv")

# Load in Kinexon Session
kinexon_session <- read.csv("data-sets/data-sets-uncompressed/data-sets-compressed/Reactive Strength Index/Kinexon Session MBB.csv")
```


## Data Cleaning


```{r clean stats dataset}
# Remove columns where all data points are NA
stats_clean <- stats |>
  select(where(~ !all(is.na(.))))


# If all three date columns are equal, remove the redundant ones
if (all(stats_clean$Date == stats_clean$Date.Reverse, na.rm = TRUE) &&
    all(stats_clean$Date == stats_clean$Session.Date, na.rm = TRUE)) {
  
  stats_clean <- stats_clean |>
    select(-c(Date.Reverse, Session.Date))
}


# Convert if Date column is NOT already in Date format
if (!inherits(stats_clean$Date, "Date")) {
  stats_clean <- stats_clean %>%
    mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
}
# Confirm it's now in Date format
class(stats_clean$Date)

stats_clean <- stats_clean[!is.na(stats_clean$Plus..Minus),]

# Add more stats
stats_clean <- stats_clean %>%
  filter(Date >= as.Date("2024-09-20", '%Y-%m-%d')) %>%
  group_by(anon_id) %>%
  mutate(Player.Average.Rebounds = mean(na.omit(Rebounds)),
         Player.Average.Steals = mean(na.omit(Steals)),
         Player.Average.Assists = mean(na.omit(Assists)),
         Player.Average.Turnovers = mean(na.omit(Turnovers)),
         Player.Average.Points.Scored = mean(na.omit(Points.Scored)),
         Player.Median.Rebounds = median(na.omit(Rebounds)),
         Player.Median.Steals = median(na.omit(Steals)),
         Player.Median.Assists = median(na.omit(Assists)),
         Player.Median.Turnovers = median(na.omit(Turnovers)),
         Player.Median.Points.Scored = median(na.omit(Points.Scored))) %>%
  ungroup() %>%
  group_by(Date) %>%
  mutate(Team.Rebounds = sum(na.omit(Rebounds)),
         Team.Steals = sum(na.omit(Steals)),
         Team.Assists = sum(na.omit(Assists)),
         Team.Turnovers = sum(na.omit(Turnovers)),
         Team.Points.Scored = as.numeric(str_split(Game.Score, "-")[[1]][1])) %>%
  ungroup() %>%
  mutate(Team.Average.Rebounds = mean(Team.Rebounds),
         Team.Average.Steals = mean(Team.Steals),
         Team.Average.Assists = mean(Team.Assists),
         Team.Average.Turnovers = mean(Team.Turnovers),
         Team.Average.Points.Scored = mean(Team.Points.Scored)) %>%
  mutate(Player.Good.Game = ifelse(Rebounds >= Player.Median.Rebounds & Steals >= Player.Median.Steals & Assists >= Player.Median.Assists,1,0),
         Player.Bad.Game = ifelse(Rebounds < Player.Median.Rebounds & Steals < Player.Median.Steals & Assists < Player.Median.Assists,1,0),
         Team.Good.Game = ifelse(Team.Rebounds >= median(Team.Rebounds) & Team.Steals >= median(Team.Steals) & Team.Assists >= median(Team.Assists),1,0),
         Team.Bad.Game = ifelse(Team.Rebounds < median(Team.Rebounds) & Team.Steals < median(Team.Steals) & Team.Assists < median(Team.Assists),1,0)) %>%
  select(-c(Sport, Day.of.Week, Month, Month.with.Number, Year, Number.of.Competitions.Won, Total.Daily.Competitions.Available, Total.Competitions.Won, Total.Game.Minutes, Offensive.Rebounds, Defensive.Rebounds, Buff.Plays, Percentage.Daily.of.Competitions.Won))

```



```{r clean kinexon_ACWR dataset}
# Remove columns where all data points are NA
kinexon_ACWR_clean <- kinexon_ACWR |>
  select(where(~ !all(is.na(.))))



# If all three date columns are equal, remove the redundant ones
if (all(kinexon_ACWR_clean$Date == kinexon_ACWR_clean$Date.Reverse, na.rm = TRUE) &&
    all(kinexon_ACWR_clean$Date == kinexon_ACWR_clean$Session.Date, na.rm = TRUE)) {
  
  kinexon_ACWR_clean <- kinexon_ACWR_clean |>
    select(-c(Date.Reverse, Session.Date))
}


# Convert if Date column is NOT already in Date format
if (!inherits(kinexon_ACWR_clean$Date, "Date")) {
  kinexon_ACWR_clean <- kinexon_ACWR_clean %>%
    mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
}
# Confirm it's now in Date format
class(kinexon_ACWR_clean$Date)


# Selecting only relevant columns
kinexon_ACWR_clean <- kinexon_ACWR_clean %>%
  select(anon_id, Date, Day.of.the.Week, Month, Year, Position, Session.Type, Session.Description, Practice.Classification, Last.7.Day.Acceleration.Load, Today.s.Acceleration.Load, Acceleration.Load.ACWR, Acceleration.Load.EWMA.ACWR, Acute.Total.Acceleration.Load, Acute.Total.Acceleration.Load.EWMA, Chronic.Total.Acceleration.Load, Chronic.Total.Acceleration.Load.EWMA, Total.Low.Acceleration.Load, Total.Medium.Acceleration.Load, Total.High.Acceleration.Load, Total.Very.High.Acceleration.Load, Last.7.Day.Total.Jump.Load, Today.s.Jump.Load, Total.Jump.Load.ACWR, Acute.Total.Jump.Load, Acute.Total.Jump.Load.EWMA, Chronic.Total.Jump.Load, Chronic.Total.Jump.Load.EWMA, Last.7.Day.Total.Minutes, Today.s.Minutes, Minutes.ACWR, Minutes.EWMA.ACWR, Acute.Minutes, Acute.Minutes.EWMA, Chronic.Minutes, Chronic.Minutes.EWMA, Total.Distance.EWMA.ACWR)
```

```{r clean the VALD_dynamo data set}
#selecting important variables, removing rows with missing data, only selecting hand movements
VALD_dynamo_clean <- VALD_dynamo %>%
  filter(Body.Region == "Hand") %>%
  mutate(Right.Side = ifelse(Repetition.Type.Laterality=="RightSide",1,0)) %>%
  select(anon_id,	Date, Max.Force.Newtons, Avg.Force.Newtons, Max.Impulse.Newton.Seconds,  Max.Rate.Of.Force.Development.Newtons.Per.Second,	Avg.Rate.Of.Force.Development.Newtons.Per.Second,	Min.Time.To.Peak.Force.Seconds, Start.Offset.Seconds, Repetition.Duration.Seconds,	Repetition.Max.Force.Newtons,	Impulse.Newton.Seconds,	Rate.Of.Force.Development.Newtons.Per.Second,	Time.To.Peak.Force.Seconds,Right.Side) %>%
  na.omit()

# Convert if Date column is NOT already in Date format
if (!inherits(VALD_dynamo_clean$Date, "Date")) {
  VALD_dynamo_clean<- VALD_dynamo_clean%>%
    mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
}
# Confirm it's now in Date format
class(VALD_dynamo_clean$Date)


#This data set contains information from each individual hand in a strength test meaning that there are two measurements per day for each player. This makes it hard to understand in plots especially when plotted like time series data. For future reference, it might be useful to take an average of both sides or note that there is a difference in each hand and plot the left and rights sides separately.

#### Note that you have to group on id and the date in order to get a valid average of both sides for a specific day for each individual player. 
```


```{r clean the ForceDecks dataset}
VALD_ForceDecks_clean <- VALD_ForceDecks %>%
  select(where(~ !all(is.na(.)))) %>%
  mutate(X=1:3384) %>%
  filter(Test.Type == "DJ") %>%
  select(X,anon_id, Date, Week, Position, Test.Type, Weight..kg., Trial, Flight.Time..s., Jump.Height..Flight.Time...cm., Eccentric.Concentric.Mean.Force.Ratio, Contact.Time..s., RSI..JH..Flight.Time..Contact.Time., RSI..Flight.Time.Contact.Time., Drop.Height..cm.)%>%
  mutate(RSI..m.per.s. = (Jump.Height..Flight.Time...cm./100)/Contact.Time..s.,
         Date = as.Date(Date, '%m/%d/%Y')) %>%
  group_by(Date, anon_id) %>%
  mutate(Weight..kg. = mean(na.omit(Weight..kg.)),
         Player.Average.RSI..m.per.s. = mean(RSI..m.per.s.)) %>%
  ungroup() %>%
  group_by(Date) %>%
  mutate(Team.Average.RSI..m.per.s. = mean(na.omit(RSI..m.per.s.)))
```


```{r kinexon session mbb cleaning}
# Remove columns where all data points are NA
kinexon_session_clean <- kinexon_session |>
  select(where(~ !all(is.na(.))))


# Convert if Date column is NOT already in Date format
if (!inherits(kinexon_session_clean$Date, "Date")) {
  kinexon_session_clean <- kinexon_session_clean %>%
    mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
}
# Confirm it's now in Date format
class(kinexon_session_clean$Date)


# Filter to only include data from this season
kinexon_session_clean <- kinexon_session_clean |>
  filter(Date >= as.Date("2024-6-01"))
```


## Recalculating Gameday Minus Columns


```{r Gameday.Minus.N.Acceleration.Load}
# Recalculating the Gameday.Minus.N.Acceleration.Load Columns

# Create a df with relevant columns and add column specifying if that day was a game or not
daily_load <- kinexon_session_clean |>
  mutate(Game.Day = if_else(Type == "Game", 1, 0)) |>
  filter(!is.na(Daily.Total.Accel.Load.Accum)) |>
  select(anon_id, Date, Daily.Total.Accel.Load.Accum, Game.Day) |>
  arrange(anon_id, Date)

# Get all game day dates per player as a lookup table
game_days <- daily_load |>
  filter(Game.Day == 1) |>
  select(anon_id, Game.Date = Date)
    

# Add column Day.Type for GD-1 - GD-4 and gamedays
daily_load <- daily_load |>
  left_join(game_days, by = "anon_id") |>
  mutate(
    Days.Until.Game = as.numeric(difftime(Game.Date, Date, units = "days")),
    Day.Type = case_when(
      Days.Until.Game == 1 ~ "GD-1",
      Days.Until.Game == 2 ~ "GD-2",
      Days.Until.Game == 3 ~ "GD-3",
      Days.Until.Game == 4 ~ "GD-4",
      Game.Day == 1 ~ "GameDay",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(Day.Type)) |>    # keep only rows within 4 days before game
  select(-Game.Date, -Days.Until.Game)

# Add columns for Gameday.Minus.N.Acceleration.Load
Gameday_Minus <- daily_load |>
  mutate(
    Gameday.Minus.1.Acceleration.Load = if_else(Day.Type == "GD-1", Daily.Total.Accel.Load.Accum, NA_real_),
    Gameday.Minus.2.Acceleration.Load = if_else(Day.Type == "GD-2", Daily.Total.Accel.Load.Accum, NA_real_),
    Gameday.Minus.3.Acceleration.Load = if_else(Day.Type == "GD-3", Daily.Total.Accel.Load.Accum, NA_real_),
    Gameday.Minus.4.Acceleration.Load = if_else(Day.Type == "GD-4", Daily.Total.Accel.Load.Accum, NA_real_),
    GameDay.Acceleration.Load = if_else(Day.Type == "GameDay", Daily.Total.Accel.Load.Accum, NA_real_)
  )

# Delete duplicate rows
Gameday_Minus <- Gameday_Minus |>
  distinct()

```


```{r Average.Gameday.Minus.N.Acceleration.Load}
# Create columns for Average.Gameday.Minus.N.Acceleration.Load

# Calculate and add columns for running averages
Gameday_Minus <- Gameday_Minus |>
  arrange(anon_id, Date) |>
  group_by(anon_id) |>
  mutate(
    # Cumulative count and sum for GD-1
    cum_count_gd1 = cumsum(!is.na(Gameday.Minus.1.Acceleration.Load)),
    cum_sum_gd1 = cumsum(replace_na(Gameday.Minus.1.Acceleration.Load, 0)),
    Average.Gameday.Minus.1.Acceleration.Load = if_else(
      cum_count_gd1 > 0,
      cum_sum_gd1 / cum_count_gd1,
      NA_real_
    ),
    
    # Cumulative count and sum for GD-2
    cum_count_gd2 = cumsum(!is.na(Gameday.Minus.2.Acceleration.Load)),
    cum_sum_gd2 = cumsum(replace_na(Gameday.Minus.2.Acceleration.Load, 0)),
    Average.Gameday.Minus.2.Acceleration.Load = if_else(
      cum_count_gd2 > 0,
      cum_sum_gd2 / cum_count_gd2,
      NA_real_
    ),
    
    # Cumulative count and sum for GD-3
    cum_count_gd3 = cumsum(!is.na(Gameday.Minus.3.Acceleration.Load)),
    cum_sum_gd3 = cumsum(replace_na(Gameday.Minus.3.Acceleration.Load, 0)),
    Average.Gameday.Minus.3.Acceleration.Load = if_else(
      cum_count_gd3 > 0,
      cum_sum_gd3 / cum_count_gd3,
      NA_real_
    ),
    
    # Cumulative count and sum for GD-4
    cum_count_gd4 = cumsum(!is.na(Gameday.Minus.4.Acceleration.Load)),
    cum_sum_gd4 = cumsum(replace_na(Gameday.Minus.4.Acceleration.Load, 0)),
    Average.Gameday.Minus.4.Acceleration.Load = if_else(
      cum_count_gd4 > 0,
      cum_sum_gd4 / cum_count_gd4,
      NA_real_
    )
  ) |>
  ungroup() |>
  select(-starts_with("cum_count_"), -starts_with("cum_sum_")) # clean up columns

```


```{r Average.Gameday.Acceleration.Load}
# Calculate running averages of Gameday.Acceleration.Load and add column Average.Gameday.Acceleration.Load

Gameday_Minus <- Gameday_Minus %>%
  arrange(anon_id, Date) %>%
  group_by(anon_id) %>%
  mutate(
    cum_count_gd = cumsum(!is.na(GameDay.Acceleration.Load)),
    cum_sum_gd = cumsum(replace_na(GameDay.Acceleration.Load, 0)),
    Average.GameDay.Acceleration.Load = if_else(
      cum_count_gd > 0,
      cum_sum_gd / cum_count_gd,
      NA_real_
    )
  ) %>%
  ungroup() %>%
  select(-cum_count_gd, -cum_sum_gd) # clean up columns

```



##Analysis

### Question 1
#### Are changes in RSI related to team game performance?

##### Team Wins and Losses
```{r Looking only at most recent season}
#filter ForceDecks data set into only the most recent season
current_season <- VALD_ForceDecks_clean %>%
  filter(Date >= as.Date("2024-09-20", '%Y-%m-%d'))

#looking at the average RSI as well as the adjusted variance
team_mean <- mean(current_season$Team.Average.RSI..m.per.s.)
team_var <- as.numeric(sample_mean_and_variance(current_season$Team.Average.RSI..m.per.s.)[2])

#looking at the team average RSI over the course of the season, the middle horizontal line represents the mean and the gold lines represent one standard deviation away from the mean
ggplot(data = current_season ,aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line(linewidth = 0.75) +
  geom_hline(yintercept = team_mean) +
  geom_hline(yintercept = team_mean + sqrt(team_var), color="#CFB87C") +
  geom_hline(yintercept = team_mean - sqrt(team_var), color="#CFB87C")

#looking at the adjusted correlations of the team averages to see if there is any correlation with time
cor(x = current_season$X, y = current_season$Team.Average.RSI..m.per.s., method = "kendall")
cor(x = current_season$X, y = current_season$Team.Average.RSI..m.per.s., method = "spearman")
```
The plot above shows  the team's average RSI per week throughout the 2024-2025 season. This plot shows that there are large variations in team's RSI values throughout the season but there does not seem to be any statistically significant underlying trend in the data. The lines in the plot represent the mean and one standard deviation from the mean throughout the season. As you can see, there is a lot of variability throughout the data. The patterns that appear though could potentially suggest an underlying ARIMA or other model for the data. 

```{r Calculating RSI changes throughout the season}

#calculating mean RSI values for each day and attaching them to correct dates
Team.means <- unique(current_season$Team.Average.RSI..m.per.s.)
Dates <- unique(current_season$Date)
IDs <- unique(current_season$anon_id)
Team.Change <- rep(0, 25)
current_season$Change.Team.Average.RSI..m.per.s. <- rep(NA, 661)

#calculating the difference in RSI from the previous measurement and attaching as a new column in the current season data set
for(i in 2:25){
  Team.Change[i] = Team.means[i] - Team.means[i+1]
  current_season[current_season$Date == Dates[i],19]<- Team.Change[i]
}

#making binary column of whether RSI increased or not
current_season <- current_season %>%
  mutate(Team.RSI.Increase = ifelse(Change.Team.Average.RSI..m.per.s. > 0, 1, 0))


#calculating the difference in RSI from the previous measurement for each player and attaching as a new column in the current season data set
current_season$Change.Player.Average.RSI..m.per.s. <- rep(0,661) #making new column

for(i in 1:16){ #cycle through players, calculate difference from last measurement
  Dates <- unique(current_season[current_season$anon_id == IDs[i],]$Date)
  Player.Means <- unique(current_season[current_season$anon_id == IDs[i],]$Player.Average.RSI..m.per.s.)
  Player.Change <- rep(0,length(Player.Means))
  ifelse(i != 14, Player.Change <- rep(0,length(Player.Means)), 0)

  for(j in 1:length(Player.Change)){
    Player.Change[j] <- Player.Means[j] - Player.Means[j+1]
    }
  #attaching in new column  
  for(k in 1:length(Player.Change)){
    current_season[current_season$Date == Dates[k] & current_season$anon_id == IDs[i],]$Change.Player.Average.RSI..m.per.s. = Player.Change[k]
  }
}

#making indicator variable for whether RSI increased from last measurement or not
current_season <- current_season %>%
  mutate(Player.RSI.Increase = ifelse(Change.Player.Average.RSI..m.per.s. > 0, 1, 0))
```


```{r Dataset for season wins and losses}
#represents games throughout the season, will merge into one data set
Win <- c(1,1,1,1,1,0,1,0,1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,1,0,0,0,1,1,1,0,0)
All_Dates <- as.Date(c("2024-10-19","2024-11-4", "2024-11-8", "2024-11-13", "2024-11-17", "2024-11-25","2024-11-26", "2024-11-27", "2024-12-2", "2024-12-7", "2024-12-13", "2024-12-21", "2024-12-30", "2025-1-4", "2025-1-8", "2025-1-12", "2025-1-15", "2025-1-18", "2025-1-21", "2025-1-25", "2025-1-28", "2025-2-2", "2025-2-5", "2025-2-8", "2025-2-11", "2025-2-15", "2025-2-18", "2025-2-22", "2025-2-24", "2025-3-2", "2025-3-5", "2025-3-8", "2025-3-11","2025-3-12", "2025-3-13", "2025-4-1"),"%Y-%m-%d")
Regular <- as.Date(c("2024-10-19","2024-11-4", "2024-11-8", "2024-11-13", "2024-11-17", "2024-12-2", "2024-12-7", "2024-12-13", "2024-12-21", "2024-12-30", "2025-1-4", "2025-1-8", "2025-1-12", "2025-1-15", "2025-1-18", "2025-1-21", "2025-1-25", "2025-1-28", "2025-2-2", "2025-2-5", "2025-2-8", "2025-2-11", "2025-2-15", "2025-2-18", "2025-2-22", "2025-2-24", "2025-3-2", "2025-3-5", "2025-3-8"),"%Y-%m-%d")
Tournament <- as.Date(c("2024-11-25","2024-11-26", "2024-11-27"),"%Y-%m-%d")
Post_Season <- as.Date(c("2025-3-11","2025-3-12", "2025-3-13", "2025-4-1"),"%Y-%m-%d")

#summarizes all games throughout the season
record <- data.frame(Date = All_Dates, Win = Win) %>%
  mutate(Type = case_when(
    Date %in% Regular ~ "Regular",
    Date %in% Tournament ~ "Tournament",
    Date %in% Post_Season ~ "Post-Season",
    TRUE ~ "NA"
  ))
```

```{r Plotting game outcomes throughout the season with average team RSI}
#plotting regular game outcomes against RSI values
ggplot(current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line(linewidth = 0.75) +
  geom_vline(xintercept = record$Date[record$Type == "Regular"], linewidth = 0.2, color = ifelse(record$Win[record$Type == "Regular"] == 1, "palegreen3", "red3"))

#plotting tournaments or post season games with average team RSI values
ggplot(current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line(linewidth = 0.75) +
  geom_vline(xintercept = record$Date[record$Type %in% c("Tournament","Post-Season")], linewidth = 0.2, color = ifelse(record$Win[record$Type %in% c("Tournament", "Post-Season")] == 1, "palegreen3", "red3"))
```

These plots show that there may not be much predictive power for team game performance from team RSI values themselves. Instead, it may be useful to look at the changes in RSI before each game or look at a normalized plot of the team's average RSI to get a look at what it would look like without as much noise. 


The data was normalized with a moving average process which removed some of the noise present in the data. For this plot, the green lines represent a game in which the team's stats were higher than the median for steals, assists, and rebounds. Red lines represent games where the total team sum of steals, assists, and rebounds are lower than the season median. Again, There does not seem to be a logical association between team average RSI and in game statistics. Instead, we find that games that had better in game performance are associated with lower RSI recordings before the game and games that had worse in game performance are associated with higher RSI recording just before the game. It's important to note that there gaps between measurements that could have missed important spikes or dips in RSI for the team that would give us a more clear understanding of the relationship. Previous research that was able to identify significant relationships between in game performance and RSI measurements were able to collect it before and after each event. 


```{r Plotting the changes in RSI throughout the season}
#plotting the change in team average RSI throughout the season
ggplot(data = current_season, aes(x=Date, y=Change.Team.Average.RSI..m.per.s.)) +
  geom_point() +
  geom_hline(yintercept = 0)


#plotting the changes in team average RSI throughout the season with important games and their outcomes
ggplot(data = current_season, aes(x=Date, y=Change.Team.Average.RSI..m.per.s.)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = stats_clean$Date[stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1], color = ifelse(stats_clean[(stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1),]$Team.Good.Game == 1, "palegreen3","red3"))
#  geom_text(data = stats_clean, x = Date, y = 0.19, label = Date, angle = 90,size = 2.75)
```
This plot shows the change in team average RSI measurements throughout the season. Each data point represents how much the average RSI changed from the previous measurement. This plot helps to make sense of how the RSI of the team changed throughout the season. We can see that while increases were more frequent, they were smaller than the drops in RSI that occurred throughout the season. 
With this though, we can see that there is potentially a relationship between changes in RSI and in team game performance. This is because visually, it seems that games where the team had a "good" game tend to have an increase in RSI measured right before the game and games that are "bad" tend to have a decrease in RSI right before the game or a less dramatic increase in RSI than games that are "good". 

```{r plotting the changes in RSI with wins and losses throughout the season}
ggplot(data = current_season, aes(x=Date, y=Team.Average.RSI..m.per.s.)) +
  geom_line(aes(color = RSI.Increase)) +
  geom_vline(xintercept = record$Date[record$Win == 1], linewidth = 0.2, color = "palegreen3")

ggplot(data = current_season, aes(x=Date, y=Team.Average.RSI..m.per.s.)) +
  geom_line(aes(color = RSI.Increase), linewidth = 0.75) +
  geom_vline(xintercept = record$Date[record$Win == 0], linewidth = 0.2, color = "red3")
```


##### Team Explosiveness Metrics
```{r plotting team scores against the team median score}
ggplot(data = current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date, color = ifelse(stats_clean$Team.Points.Scored > median(stats_clean$Team.Points.Scored), "palegreen3", "red3"), linewidth = 0.2)

```
This plot represents if the team scored higher or lower than the median team score in that game. Green represents higher and red represents lower than the median. This plot suggests that lower than median scores are associated with drops in RSI before the game (found in the middle of the season). As the season goes on, we can see that even though the team's RSI goes up, the team has a consistently lower than median score. This could be indicative of trouble recovering from fatigue found in earlier research.

```{r Looking at team summary statistics compared to median with RSI values throughout the season}
#Looking at games that had team total rebounds above or below median team season value
ggplot(data = current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date, color = ifelse(stats_clean$Team.Rebounds > median(stats_clean$Team.Rebounds), "palegreen3", "red3"), linewidth = 0.2)

#Looking at games that had team total steals above or below median team season value
ggplot(data = current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date, color = ifelse(stats_clean$Team.Steals > median(stats_clean$Team.Steals), "palegreen3", "red3"), linewidth = 0.2)

#Looking at games that had team total assists above or below median team season value
ggplot(data = current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date, color = ifelse(stats_clean$Team.Assists > median(stats_clean$Team.Assists), "palegreen3", "red3"), linewidth = 0.2)
```

```{r Plotting games that are considered overall good or bad throughout season}
ggplot(data = current_season, aes(Date, Team.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date[stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1], color = ifelse(stats_clean[(stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1),]$Team.Good.Game == 1, "palegreen3","red3"))
```

###### Applying regularization techniques to team data
```{r}
#smooth <- exponential_smoothing(unique(current_season$Team.Average.RSI..m.per.s.),2,0)
ma <- moving_average(unique(current_season$Team.Average.RSI..m.per.s.), 1)
#x53 <- five_three_x(unique(current_season$Team.Average.RSI..m.per.s.))

#current_season$Team.Average.RSI.Smooth <- rep(NA,661)
current_season$Team.Average.RSI.Moving.Average <- rep(NA, 661)
#current_season$Team.Average.RSI.53X <- rep(NA, 661)

for(i in 1:25){
  #current_season[current_season$Date == Dates[i],21]<- smooth[i]
  current_season[current_season$Date == Dates[i],22]<- ma[i]
  #current_season[current_season$Date == Dates[i],23]<- x53[i]
}

ggplot(data=current_season, aes(Date, Team.Average.RSI.Moving.Average)) +
  geom_line() +
geom_vline(xintercept = stats_clean$Date[stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1], color = ifelse(stats_clean[(stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1),]$Team.Good.Game == 1, "palegreen3","red3"))
```

###### Assessing effectiveness of RSI as predictor of in game success
```{r}
Game.Days <- unique(stats_clean$Date)
Game.Number <- seq(30, 1, -1)

for(i in 1:30){
  stats_clean[stats_clean$Date == Game.Days[i],]$Game.Number <- Game.Number[i]
}

```







### Question 2
#### Are changes in RSI related to individual statistical game performance?

```{r Looking at "ID_42" for this first analysis}
#Looking at games that had above or below median rebounds throughout the season
ggplot(data = current_season[current_season$anon_id == "ID_42",], aes(Date, Daily.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date[stats_clean$anon_id == "ID_42"], color = ifelse(stats_clean[stats_clean$anon_id == "ID_42",]$Rebounds >= median(stats_clean[stats_clean$anon_id == "ID_42",]$Rebounds), "palegreen3", "red3"))

#Looking at games that had above or below median steals throughout the season
ggplot(data = current_season[current_season$anon_id == "ID_42",], aes(Date, Daily.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date[stats_clean$anon_id == "ID_42"], color = ifelse(stats_clean[stats_clean$anon_id == "ID_42",]$Steals >= median(stats_clean[stats_clean$anon_id == "ID_42",]$Steals), "palegreen3", "red3"))

#Looking at games that had above or below median assists throughout the season
ggplot(data = current_season[current_season$anon_id == "ID_42",], aes(Date, Daily.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date[stats_clean$anon_id == "ID_42"], color = ifelse(stats_clean[stats_clean$anon_id == "ID_42",]$Assists >= median(stats_clean[stats_clean$anon_id == "ID_42",]$Assists), "palegreen3", "red3"))
```



```{r Looking at overall good or bad game for "ID_42"}
ggplot(data = current_season[current_season$anon_id == "ID_42",], aes(Date, Daily.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date[stats_clean$anon_id == "ID_42" & (stats_clean$Player.Good.Game == 1 | stats_clean$Player.Bad.Game ==1)], color = ifelse(stats_clean[stats_clean$anon_id == "ID_42" & (stats_clean$Player.Good.Game == 1 | stats_clean$Player.Bad.Game ==1),]$Player.Good.Game == 1, "palegreen3","red3"))
```

```{r}
ggplot(data = current_season[current_season$anon_id == "ID_61",], aes(Date, Daily.Average.RSI..m.per.s.)) +
  geom_line() +
  geom_vline(xintercept = stats_clean$Date[stats_clean$anon_id == "ID_61" & (stats_clean$Player.Good.Game == 1 | stats_clean$Player.Bad.Game ==1)], color = ifelse(stats_clean[stats_clean$anon_id == "ID_61" & (stats_clean$Player.Good.Game == 1 | stats_clean$Player.Bad.Game ==1),]$Player.Good.Game == 1, "palegreen3","red3"))
```


#### applying regularization techniques to team data
```{r}
smooth <- exponential_smoothing(unique(current_season$Team.Average.RSI..m.per.s.),2,0)
ma <- moving_average(unique(current_season$Team.Average.RSI..m.per.s.), 1)
x53 <- five_three_x(unique(current_season$Team.Average.RSI..m.per.s.))

current_season$Team.Average.RSI.Smooth <- rep(NA,661)
current_season$Team.Average.RSI.Moving.Average <- rep(NA, 661)
current_season$Team.Average.RSI.53X <- rep(NA, 661)

for(i in 1:25){
  current_season[current_season$Date == Dates[i],21]<- smooth[i]
  current_season[current_season$Date == Dates[i],22]<- ma[i]
  current_season[current_season$Date == Dates[i],23]<- x53[i]
}

ggplot(data=current_season, aes(Date, Team.Average.RSI.Moving.Average)) +
  geom_line() +
geom_vline(xintercept = stats_clean$Date[stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1], color = ifelse(stats_clean[(stats_clean$Team.Good.Game ==1 | stats_clean$Team.Bad.Game ==1),]$Team.Good.Game == 1, "palegreen3","red3"))
```


### Question 3
#### Is the previous week's load related to RSI?


### Question 4
#### What is each athlete's variation in RSI? What is a meaningful changes in RSI for the team and for individual athletes?